\section{Analysis of the Encounter Protocol}
\label{sectionanalysis}

In this section, we prove that with a clique of $k$ agents,  
our protocol guarantees each agent can record all
the peers in $O(k)$ slots with high probability.
Note that, $k$ is not known in the execution of the protocol but just for analysis
in this section.
Formally, this result is derived from the following theorem.
\begin{theorem}
    \label{bound}
    Consider an encounter process of $k$ agents~($k$ is a integer $k \geq 2)$ :
    \begin{itemize} 
    \item[(1)] An agent in detecting stage  will switch to the connecting stage in $O(\theta^{-2})$
    slots with high probability. Recall that $\theta$ is the pre-defined duty cycle and can be seen 
    as a constant.
    \item[(2)] When all the agents are in the connecting stage, all of them will
    transmit successfully in a first sub-slot and turn to the \emph{quiet} state 
    in $O(k)$ slots with high probability.
    \end{itemize}
\end{theorem}

We prove these two conclusions in Theorem~\ref{bound} in Section~\ref{boundSW}, and
Section~\ref{boundPC} respectively. 
Based on this theorem, we can
estimate a duration of the encounter process
between any pair of agents $u_i$ and $u_j$, 
by checking the records in their local logs.  
If there are $z$ consecutive records of the other peer, it indicates a 
duration of $2z\hat{T}t_0$. Recall that $2t_0$ is the real time 
of a slot and $\hat{T}$ is the number of slots of a round in 
the connecting stage, which is 
sufficiently large to satisfy the bound 
of {\pName}  while still short enough relative to the 
time required for an agent to move a short distance. 

\subsection{Time bound for detecting stage}
\label{boundSW}

When an encounter happens, an agent in detecting stage will switch to connecting stage very soon. 
We derive this conclusion from the following three lemmas.

\begin{lemma}
    \label{RDS}
    Consider any two agents $u_i$ and $u_j$ in detecting stage.
    In each period $T_0$, they will turn on the radio in the same slot at least once.
\end{lemma}
\begin{IEEEproof}
Assume the time drift between $u_i$ and $u_j$ is $\delta_{ij}$ (mod $T_0$).
In the RDS constructed under $T_0$, there exists at least one ordered pair $(a_i,a_j)$ such that $a_i - a_j \equiv \delta_{ij}$ (mod $T_0$).
Thus the $a_i^{th}$ slot in a period of $u_i$ is exactly the $a_j^{th}$ slot in a period of $a_j^{th}$ and both of them
turn on the radio in this slot according to Alg.~\ref{DA} Line~\ref{On}, which completes the proof.
\end{IEEEproof}

\begin{lemma}
    \label{dtoc}
    Consider $k$ agents all in the detecting stage at the beginning, 
    an agent in detecting stage will turn to the connecting stage in 
    $O(\theta^{-2})$ slots with high probability.
\end{lemma}
\begin{IEEEproof}
    By Lemma~\ref{RDS}, an agent can turn on the radio in the same 
    slot with any other peer at least once during a period of $T_0$.
    The probability it detects a peer in a period of $T_0$ is at least
    $Pr \geq 1 - \frac{1}{2^{k-1}}$.
    Hence given a small enough constant $\eta$,  
    it holds with high probability that an agent will turn to connecting stage in 
    $\frac{\ln\eta}{\ln\frac{1}{2^{k-1}}}$ periods, which is 
    $\frac{\ln\eta}{\ln\frac{1}{2^{k-1}}} \lceil \frac{9}{4\theta^{2}} \rceil$ slots
    in total.    
\end{IEEEproof}
\vspace{0.1in}
\begin{lemma}
    \label{case}
    An agent turns to the connecting stage will increase the probability that other
    peers in the detecting stage to detect peers.
\end{lemma}
\begin{IEEEproof}
    An agent in the detecting stage will transmit or listen in every first sub-slot 
    which can help peers to detect it,
    while it only turns on the radio a fraction of time in the detecting stage.
\end{IEEEproof}
   
According to Lemma~\ref{dtoc} and \ref{case}, every agent will turn to the connecting stage in 
$O(\theta^{-2})$ slots with high probability. 

\subsection{Time bound for connecting stage}
\label{boundPC}
Consider the $k$ agents all in the connecting stage.
We analyze the upper bound 
for all the agents in the clique to transmit successfully.
Note that, an agent transmitting successfully will be recorded by 
all the other peers in the clique, and then it switches to the \emph{quiet} state. 

Denote $S_t$ as the set of agents
which have not switched to the \emph{quiet} state in slot $t$, and 
$|S_t|$ is the cardinality. 
Thus the upper bound of the protocol is the maximum slots
for a clique to turn to $|S_t| = 0$.

In the following, we first show the
upper bound of slots it takes from the beginning to $|S_t| \leq \log k$ in Lemma~\ref{fast}.
Then we show the upper bound it takes to $|S_t| = 1$ in Lemma~\ref{slow}.
Finally we present that it takes for all agents to successfully transmit~(i.e, $|S_t| = 0$.) in $O(k)$ 
in Lemma~\ref{one}. 

\begin{lemma}
    \label{fast}
    There exists $T_2=O(k)$ it holds with high probability that $|S_{T_2}| \leq \log k$. 
\end{lemma}

To prove this lemma, we need to introduce and prove some small lemmas at first.
We review two useful lemmas as follows.
\begin{lemma}
    \label{sumP}
    Consider a set of $l$ agents, $u_1, u_2,\dots, u_l$.
    For an agent $u_i$, it transmits with probability 
    $0 < \omega(u_i) < \frac{1}{2}$. Let $p_0$ denote the 
    probability that the channel is idle in a slot; and $p_1$ denote 
    the probability that there is exactly one transmission in a slot.
    Then $p_0\cdot {\sum}_{i=1}^{l}\omega(u_i)\leq p_1 \leq
    2\cdot p_0\cdot {\sum}_{i=1}^{l}\omega(u_i)$     
\end{lemma}
\begin{lemma}
    \label{mulP}
    With $a_i \in [0, \frac{1}{2}]$ for $i = 0,1,\dots,$ it holds that
    \begin{equation}
        \label{mul}
        4^{-{\sum}_{i}a_i}\leq {\prod}_{i}(1-a_i)\leq e^{-{\sum}_{i}a_i}. 
    \end{equation}
\end{lemma}
The proof of Lemma~\ref{sumP} can be referred to~\cite{Richa2010A}
and the proof of Lemma~\ref{mulP} can be referred to~\cite{Daum2013Maximal}.
Based on these two lemmas, we get the following conclusion.

\begin{lemma}
    \label{safe}
    For a slot $T>0$ with $|S_t| \ge \log k$, if 
    there exist constants $\alpha_1, \alpha_2 \ge 1$ such that
    $\alpha_1 \leq \sum_{u\in S_t}{\omega}_t(u) \leq \alpha_2$, then with constant probability 
    there is one active node switching to the \emph{quiet} state
    in each slot. 
\end{lemma}
\begin{IEEEproof}
    By Lemma~\ref{mulP} and~\ref{safe}, in each slot, the channel is idle with probability at least 
    $4^{-2\alpha_2}$, and there is exactly one transmission on the channel
    with probability at least $\alpha_1\cdot 4^{-2\alpha_2}$. Thus, 
    it holds that with constant probability (given $\alpha_1$ and $\alpha_2$),
    there are one agent switch to the quiet state in each slot.
\end{IEEEproof}

Next we show that after all the agents turn to the connecting stage, 
the summation of the total agents' transmission probabilities will go
between $\alpha_1$ and $\alpha_2$ soon, where $\alpha_1$ and $\alpha_2$
are constants defined in Lemma~\ref{safe}.
\begin{lemma}
    \label{lemma7}
    For a slot with ${\sum}_{u\in S_t}\omega_t(u)=\alpha$, it holds that
    $Pr[{\sum}_{u\in S_t}\omega_{t+1}(u) \leq \alpha\cdot\frac{3}{4}] 
    \geq \frac{7}{8}$
    for large enough $\alpha$.
\end{lemma}
\begin{IEEEproof}
    The probability that there are more than one agents transmitting in slot 
    $t$ is at least $1-exp\{-\alpha\}$ according to Equation~(\ref{mul}).
    All the agents will halve their transmission probabilities if the channel is 
    not idle in slot $t$. Denote X as the random variable that indicates the value of 
    ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u)$. We get,
    \begin{equation*}
        Pr[X ={\sum}_{u\in S_t}\frac{\omega_t(u)}{2}] 
        \geq 1 - exp\{-\alpha\}
    \end{equation*}
    which is at least $7/8$ when $\alpha > 2.1$. Hence,
    \begin{equation*}
        X \leq \frac{7}{8}\cdot\frac{1}{2}\cdot\alpha + 
        \frac{1}{8}\cdot 2\alpha 
        < \frac{3}{4}\cdot\alpha.
    \end{equation*}
    Therefore, it holds with high probability that for large $\alpha$, $Pr[{\sum}_{u\in S_t}\omega_{t+1}(u) 
    \leq \alpha\cdot\frac{3}{4}] \geq \frac{7}{8}$.
    % Note that, 
    % we did not consider the effect when an agent turns to the \emph{quiet} state,
    % which only makes the summation decrease and hence is not harmful. 
\end{IEEEproof}

\begin{lemma}
    \label{lemma8}
    There exists a constant $\hat{\alpha_2} > 1$,
    such that among $\gamma\log k$ slots (not necessarily consecutive)
    with ${\sum}_{u\in S_t}\omega_t(u) \geq \hat{\alpha_2}$ and sufficiently
    large $\gamma > 0$, there are at least $\frac{3}{4}\gamma\log k$ slots
    with ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u) < \frac{3}{4}{\sum}_{u\in S_t}\omega_t(u)$,
    with probability $1-O(k^{-1})$.
\end{lemma}
\begin{IEEEproof}
    Let $T := \gamma\log k$, and $X_t$ be the random variable that indicates the value of 
    ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u) / {\sum}_{u\in S_t}\omega_t(u)$. Then
    by Lemma~\ref{lemma7}, it holds that $Pr[X_t \leq \frac{3}{4}] \geq \frac{7}{8}$.
    Let $Y_t$ be the binary random variable that takes value $1$ if $X_t \leq \frac{3}{4}$.
    Note that given ${\sum}_{u\in S_t}\omega_t(u) \geq \hat{\alpha_2}$, $E[Y_t] \geq \frac{7}{8}$
    always hold. Hence, $E[\sum_{t=1}^{T}Y_t] \geq T\cdot\frac{7}{8}$, and it holds that
    $Pr[\sum_{t=1}^{T}Y_t\leq T\cdot\frac{3}{4}]$ by the Chernoff bound. That is with probability 
    $1-O(k^{-1})$, there are at least $T\cdot\frac{3}{4}$ slots $t$ with 
    ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u) / {\sum}_{u\in S_t}\omega_t(u) \leq 3/4$, which completes the proof. 
\end{IEEEproof}


\begin{lemma}
    \label{lemma9}
    There exists a constant $\hat{\alpha_2} > 1$, such that
    during any period of $\gamma\log k$ slots with sufficiently large 
    $\gamma >0$, the probability that within the considered period there
    is a slot $t$ with ${\sum}_{u\in S_t}\omega_t(u) \leq \hat{\alpha_2}$
    is $1-O(k^{-1})$.
\end{lemma}
\begin{IEEEproof}
    Denote $T :=\gamma\log k$ and the period of $T$ starts from slot $t_0$.
    By Lemma~\ref{lemma8}, with probability at least $1-O{n^{-1}}$, it holds that 
    \begin{align*}
        {\sum}_{u\in S_{t_0+T}}\omega_{t_0+T}(u)  &\geq 
        {\sum}_{u\in S_{t_0}}\omega_{t_0}(u)\cdot(\frac{3}{4}\cdot\frac{3}{4}
        \cdot\frac{3}{4}\cdot 2)^{\frac{T}{4}} \\
       &={\sum}_{u\in S_{t_0}}\omega_{t_0}(u)\cdot(\frac{27}{32})^{\frac{T}{4}}.
    \end{align*} 
    Since $\sum_{u\in S_{t_0+T}}\omega_{t_0+T}(u) < k$ and $T = \gamma\log k$,
    we know that $\sum_{u\in S_{t_0+T}}\omega_{t_0+T}(u)$ is at most $\hat{\alpha_2}$ 
    for large enough $\gamma$.
\end{IEEEproof}


\begin{lemma}
    \label{lemma10}
    There exists a constant $\hat{\alpha} \geq 0.01$ such that
    for any time $t$ with ${\sum}_{u\in S_t}\omega_t(u)=\hat{\alpha}$,
    it holds that
    \begin{equation}
        Pr[{\sum}_{u\in S_{t+1}}\omega_{t+1}(u) \geq \hat{\alpha}\cdot \frac{4}{3}] \geq \frac{7}{8}.
    \end{equation}
\end{lemma}
\begin{IEEEproof}
    Denote X as the random variable that indicates the value of 
    ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u)$. The probability that 
    there are no transmissions in slot 
    $t$ is at least $4^{-\hat{\alpha}}$. Hence,
    \begin{equation*}
        Pr[X ={\sum}_{u\in S_t}2\cdot\omega_t(u)] 
        \geq 1 - 4^{-\hat{\alpha}}
    \end{equation*}
    which is at least $7/8$ when $\hat{\alpha}$ is close to $0.01$. Hence,
    \begin{equation*}
        X \geq \frac{7}{8}\cdot 2\cdot\hat{\alpha} + 
        \frac{1}{8}\cdot \frac{1}{2}\hat{\alpha} 
        > \frac{4}{3}\cdot\hat{\alpha}.
    \end{equation*}
    Therefore, it holds with high probability that for small $\hat{\alpha}$ close to $0.01$, $Pr[{\sum}_{u\in S_t}\omega_{t+1}(u) 
    \leq \hat{\alpha}\cdot\frac{4}{3}] \geq \frac{7}{8}$.


\end{IEEEproof}


\begin{lemma}
    \label{lemma11}
    There exists a constant $\hat{\alpha_1} > 0$, such that among $\gamma\log k$ 
    slots~(not necessarily consecutive) with ${\sum}_{u\in S_t}\omega_t(u)
    \leq \hat{\alpha_1}$ and sufficiently large $\gamma > 0$, there are at 
    least $\frac{3}{4}\gamma\log k$ slots with ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u)
    \geq \frac{4}{3}{\sum}_{u\in S_t}\omega_t(u)$, with probability $1 - O(k^{-1})$.
\end{lemma}
\begin{IEEEproof}
    Denote $T := \gamma\log k$, $X_t$ as the random variable that indicates the value of 
    ${\sum}_{u\in S_{t+1}}\omega_{t+1}(u) / {\sum}_{u\in S_t}\omega_t(u)$, and 
    $Y_t$ to be the binary random variable that takes value $1$ if $X_t \leq \frac{4}{3}$.
    Note that given ${\sum}_{u\in S_t}\omega_t(u) \geq \hat{\alpha_1}$, $E[Y_t] \geq \frac{7}{8}$
    always hold. Hence, $E[\sum_{t=1}^{T}Y_t] \geq T\cdot\frac{7}{8}$, and it holds that
    $Pr[\sum_{t=1}^{T}Y_t\leq T\cdot\frac{3}{4}]$ by the Chernoff bound. That is with probability 
    $1-O(k^{-1})$, there are at least $T\cdot\frac{4}{3}$ slots $t$ with  
\end{IEEEproof}


\begin{lemma}
    \label{lemma12}
    Let $t_0$ be the first slot in which ${\sum}_{u\in S_t}\omega_t(u)$
    drops below $\hat{\alpha_2}$. in the subsequent $T :=\tau\cdot\log k$ slots
    where $\tau > 0$ and $k$ is large enough, the following hold:
    \begin{itemize}
        \item[(1)] there are at least $\frac{3}{4}\cdot T$ slots $t$
        with ${\sum}_{u\in S_t}\omega_t(u) \leq \alpha_2$, where $\alpha_2 > \hat{\alpha_2}$
        is a constant.
        \item[(2)] there are at least $\frac{3}{4}\cdot T$ slots $t$
        with ${\sum}_{u\in S_t}\omega_t(u) \geq \alpha_1\cdot k$, where $\alpha_1 < \hat{\alpha_1}$
        is a constant.
    \end{itemize}
\end{lemma}
% \begin{IEEEproof}
%     First we prove the conclusion (1).

% \end{IEEEproof}

With Lemma~\ref{safe},~\ref{lemma7},~\ref{lemma8},~\ref{lemma9},~\ref{lemma10},~\ref{lemma11} and
\ref{lemma12}, we now prove Lemma~\ref{fast}.
% \begin{IEEEproof}
%     By Lemma~\ref{safe}, 
% \end{IEEEproof}


Then we show that when the number of agents not switching to 
the \emph{quiet} state is less than $\log k$, the process accelerates to 
only one agent not in \emph{quiet} state with high probability.
This is formulated as the following lemma.

\begin{lemma}
    \label{slow}
    At time $T_3 :=T_2 +\gamma_2 \cdot \log k = O(k)$ it holds with high
    probability that $|S_{T_3}| = 1$.
\end{lemma}
\begin{IEEEproof}
    After $T_2$ slots, it holds with high probability 
    that $|S_{T_2}| \leq \log k$. Then it takes at most $\gamma_2 \cdot \log k$
    slots to keep $\beta_1 \leq \sum_{u\in S_t} \omega_t^u \leq \beta_2$,
    where $\beta_1$ and $\beta_2$ are two constants.
    Afterwards, there is a slot $T_3 :=T_2 +\gamma_2\cdot\log k$ such that
    $|S_{T_3}| = 1$, where $\gamma_2 > 0$ is a large enough constant. 
    Otherwise during the period from $T_2$ to $T_3$,
    with high probability there are 
    more than $\log k$ agents switching to
    the \emph{quiet} state, which can not be possible.
\end{IEEEproof}

Finally we need to prove the following lemma.
\begin{lemma}
    \label{one}
    For time $T_3$ when there is only one agent not switching to \emph{quiet} state,
    at time $T_4 := 2 \cdot T_3 + O(1) = O(k)$ it holds with high 
    probability that this agent successes to transmit.
\end{lemma}
\begin{IEEEproof}
    Since at time $T_3$ there is only one agent still attempting to transmit,
    the transmission probability will get to $\zeta$ at time $2\cdot T_3$ if it keeps
    listening. Note that if agent $u$ transmit with probability $\zeta$ for $\log k$ slots,
    where $d$
    then with high probability there exist one slot in which agent $u$ successes to transmit.
\end{IEEEproof}

According to the Lemma~\ref{fast}, \ref{slow} and \ref{one},
we get the upper bound $O(k)$ of {\pName} with respect to
the clique size $k$. Note that, since the round of each agent may not be 
synchronized, {\pName} requires an overlap of at least $T_4$ time 
slots in a complete round between each agent.
This can be easily solved as we set $\hat{T} := 2\cdot T_4$ in Algorithm~\ref{CA}. 




