\section{Introduction}

Collecting detailed information about wild animals remains a significant technical challenge that
limits the ability of Ecologists to study them, the interactions between them, and the interactions 
between wild animals and their environment. One tools that has emerged a little over a decade ago
and has been gaining significance is {\em encounter detection and logging}.

Encounter detection systems consist of radio devices called {\em tags} that are attached to wild
animals (and sometimes also to fixed positions and to livestock). The radios transmit identification
packets periodically, usually once every few seconds. The radios also listen to such packets, not
necessarily continuously, and record data about received packets in persistent memory on the tag. The
radios are typically configured for short-range communication by using low transmit power and by 
using high data-rates (both limit the signal-to-noise ratio at the receiver). Because the radios
are configured for short-range communication, receiving a packet implies that the transmitting tag is
in close proximity to the receiving tag. Recent systems record with each packet a received signal strentgh
indication (RSSI), which helps to estimate the distance between the transmitter and receiver.
This is the main goal of these systems: to log close-proximity
events between two or more animals. The logs are downloaded either by physically retrieving the tags,
or remote download to base-stations placed in locations that the animals are known to frequent.

Such systems have gained popularity among Ecologists because the tags are relatively inexpensive and can
be very small, because deployment does not require much infrastructure in the field, and because close-proximity
encounters are key aspects of many significant events in the life of animals: mating, predation, transmission of infectious
diseases, and so on. We comment on some of these applications in Section~\ref{sec:related-work}.

Encounter-registration systems pose challenging protocol-ralated problems. One problem is to minimize
the power used for listening to other tags. All of the systems developed so far have used low-power integrated
UHF transceivers, and the power consumed by these while receiving is comparable to the power that they use
to transmit. The dominant strategy to address this issue appears to be duty-cycling the receiver, so that it is
off most of the time. Synchronizing the timing of the activity periods of all tags can minimize total receive energy
while maximizing that transmissions will be received. If tags do not synchronize their activity periods, many transmissions
at close proximity may go unnoticed; this is not an unreasonable solution if encounters are long. Another emerging approach
to the receive-power problem is based on so-called {\em wake-up receivers}, specialized nano-power sub-receivers whose only
job is to wake up the main receiver when a strong packet is received. These have not been used in Ecology research so we
exclude them from most of the discussion; we do comment on them in Section~\ref{sec:related-work}.

The other protocol-related problem is interference. If many tags transmit simultaneously, receiving tags may fail
to reconstruct valid packets. Paradoxically, highly-efficient systems with good time synchronization and short activity
periods suffer more from this problem than less efficient systems in which the clocks of tags are not synchronized tightly
enough to transmit simultaneously. This is not a theoretical problem; many species of animals, including many species of
birds and bats, roost together, so encounters of tens or hundreds of individuals are not necessarily rare.

This paper proposes effective protocols for encounter registration, addressing these two problems systematically and
methodically. 

HERE WE NEED SHEN TONG TO ADD A SUMMARY OF THE TECHNICAL RESULTS (WHAT WE HAVE BEEN ABLE TO SHOW).



 

