\section{Related Work}
\label{sec:related-work}

\subsection{Encounter-registration system}
The first encounter-registration system was developed by a company called Sirtrack 
over a decade ago~\cite{Prangle2006NewRadiocolars}. The tags
are placed on collars attached to wild mammals. 
They are fairly heavy, weighing 45--450g, depending on the
size of the battery. % Tags transmit every 1.5~s, and are able to record encounters of 15~s or longer. 
This system has been commercially used, for example, to study 
the possibility of disease transmission between cattle 
and wild badgers~\cite{Bohm2009WildlifeLivestock}.

The next generalization is a system called {\em Enounternet}~\cite{Menhill2012NovelTelemetry,Rutz2012AutomatedMapping}, 
which has been available commercially for a few years. 
The key innovation of Encounternet was tags' weight: 
tags used a tiny printed-circuit board and
could be powered by miniature batteries, allowing tags weighing 1.3g and up to be manufactured. 
Clearly, the small batteries restrict the life-span of tags and increase the importance of effective protocols, 
especially with respect to low-power operation. 
The small size of Encounternet tags has enabled the study of interaction of small species, 
including small birds~\cite{Levin2015Performance} and freshwater fish~\cite{Tentelier2016FishNetwork}.
% Another innovation of the Encounternet system is the recording of RSSI with every reception report, allowing
% researchers to estimate the distance of each encounter. 

More recently, prototype Encounternet-like tags have been developed and tested on
bats~\cite{Ripperger2016ProximitySensing,dressler2016bats}. These tags include a wake-up receiver, which
enables nano-power listening without maintenance of inter-tag synchronized clocks. These tags  
do not appear to have been commercialized or widely-used.




\subsection{Protocol for encounter registration}


Unfortunately, none of the papers that describe these systems and 
ecological research give any details on the
MAC-layer protocol that is used, nor on how the receiver is duty cycled. 
Existing methods to encounter registration problem are mainly based on fixed transmitting 
probability~\cite{Menhill2012NovelTelemetry,Rutz2012AutomatedMapping}
(i.e., agents transmit a beacon with a fixed probability $p$ and listen with $1-p$). 
In particular, it appears that many of these 
protocols are not particularly efficient. For example, switching 
an Encounternet tag with a particular battery from
transmit-only mode 
to encounter-registration mode reduces the lifespan
of the tag from 7.5 days to less than a day, indicating that the 
receiver is active a significant fraction of the time.

Although to the best of our knowledge, there has been no analyses of the protocols 
for encounter registration problem,
several similar problems are well-studied in the wireless network literature, 
such as  
minimum dominating set problem~\cite{Scheideler2008An,Yu2013Review},
neighbor discovery problem~\cite{Bakht2012Searchlight, Sun2014Hello,Chen2015On}
and information exchange problem~\cite{Capetanakis1979Tree,Daum2013Maximal,Yu2017Uniform}.

Minimum dominating set problem~\cite{Scheideler2008An,Yu2013Review} is studied to 
deal with interference challenges for the wireless multi-hop networks.
This problem focus on the communication models, such as unit disk graph model~\cite{Lebhar2009Unit} (UDG),
graph-based model~\cite{De2007A} and Signal to Interference plus Noise Ratio model~\cite{Lee2007Signal} (SINR).
Neighbor discovery problem is well studied in the wireless sensor 
networks. Protocols~\cite{Bakht2012Searchlight, Sun2014Hello,Chen2015On} 
are designed for two nodes to discover each other and 
are applied directly to the multi-node scenario.
Information exchange problem~\cite{Capetanakis1979Tree,Daum2013Maximal,Yu2017Uniform} is 
studied on the information propagation in a single-hop network. 
In the network, there are active nodes with packets to transmit 
and inactive nodes waiting to receive.

However, protocols for these problems are not well applicable to encounter-registration systems.
Whether an encounter is happening at the moment is unknown, making the problem much more complicated,
and thus high efficiency (when an encounter happens)
and low-energy consumption (when it does not) 
can hardly be mutually exclusive.

% Why are the protocols for these problems not applicable to encounter registration problem?
% the key characteristic satisfy 

% A common solution to these problems is to control the transmission probability of the node in the 
% network, making it optimal to transmit successfully.
% Some deterministic methods, such as quorum system~\cite{Peleg1995The} and prime number, are also 
% adopted to help synchronization in the protocol design.
% The major difference between the problems above and the encounter registration problem
% in this paper is the mobility of the wild animals. It makes the problem 
% complicated that whether an encounter is happening at the moment is unknown.
% Besides the encounter behavior varies from individual to individual. Some animals are eager to 
% keep company with peers while others stay alone most of the time.
% Since animals are attached with energy-restrained
% tags, the energy-efficiency should also be taken into consideration.
% All these factors make the problem challenging.
  